---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE}
library(dplyr)
library(ggplot2)
activity <- read.csv("activity.csv")
summary(activity)
```


## What is mean total number of steps taken per day?
First we have to process the data before analysing it to answer the question. 

```{r echo=TRUE}
activity$date <- as.Date(activity$date)
```

Now we sum the number of steps per day using the function tapply. Next I create the histogram.
```{r echo=TRUE}
stepsday <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
hist(stepsday, breaks=10, col="darkgreen",main="Histogram of steps per day", xlab="Steps per day")
```

```{r echo=TRUE}
stepsmean <- mean(stepsday)
stepsmedian <- median(stepsday)
```
The mean of the total steps per day is `r stepsmean`. The median is `r stepsmedian`.
 


## What is the average daily activity pattern?
 
```{r echo=TRUE}
intervsteps <- tapply(activity$steps, activity$interval, mean, na.rm=TRUE)
plot(names(intervsteps), intervsteps, type="l", main = "Average number of steps per interval", ylab="Steps", xlab="Interval")

``` 

```{r echo=TRUE}
intervmax <- max(intervsteps)
nameintervmax <- names(intervsteps[intervmax])
```

The interval with the highest average number of steps is `r nameintervmax`.



## Imputing missing values

```{r echo=TRUE}

numNA <- sum(is.na(activity$steps))
```
The total number of missing (NA) values is `r numNA`.

We're going to substitute all the missing values for the mean of steps.

```{r echo=TRUE}
intervmean <- mean(activity$steps, na.rm=TRUE)
actifill <- activity
for(i in 1:nrow(actifill)) {
        if(is.na(actifill$steps[i])) {
                actifill$steps[i] <- intervmean
        }
}

stepsfilledday <- tapply(actifill$steps,actifill$date,sum)

hist(stepsfilledday, main="Steps per day", xlab="Number of steps")
```
```{r echo=TRUE}
meanfilled <- mean(stepsfilledday)
medfilled <- median(stepsfilledday)

```
The mean steps per day is now `r meanfilled`, and the median is `r medfilled`. Both values have increased and in fact now have the same value.


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}

actifill$weekday <- ifelse(weekdays(actifill$date) == "sábado" | weekdays(actifill$date) == "domingo", "weekend", "weekday")


daystep <- filter(actifill, actifill$weekday=="weekday")
daystep2 <- tapply(daystep$steps,daystep$interval,mean)

endstep <- filter(actifill, actifill$weekday=="weekend")

endstep2 <- tapply(endstep$steps,endstep$interval,mean)

par(mfrow=c(2,1))
plot(names(endstep2),endstep2, type="l", main="Mean steps per interval, Weekend", xlab="Steps", ylab="Interval")
plot(names(daystep2),daystep2, type="l", main="Mean steps per interval, Weekdays", xlab="Steps", ylab="Interval")


```



